---
title: "P8105_hw1_jr3869"
output: github_document
---


# problem 1

Create a data frame comprised of:

A random sample of size 10 from a uniform[0, 5] distribution
A logical vector indicating whether elements of the sample are greater than 2
A (length-10) character vector
A (length-10) factor vector

```{r problem1}
library(tidyverse)
set.seed(1)

vec_df = tibble(
  vec_data= runif(10, min=0, max=5),
  vec_logical = vec_data >2,
  vec_char = c("hello","world","my","name","is","Jingwei","Ren","a","b","c"),
  vec_factor=factor(c("male","male","female","female","d","e","f","g","h","i"))
)
```

### mean of each variable

```{r mean}
mean(vec_df$vec_data)
mean(vec_df$vec_logical)
mean(vec_df$vec_char)
mean(vec_df$vec_factor)
```

* mean function works for random sample and logical vector, but does not work for character and factor vector.This is because mean is only meaningful with numerical and logical vector. 

* Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output).

### convert to numeric part1 
```{r numeric, eval=FALSE}
as.numeric(vec_df$vec_logical)
as.numeric(vec_df$vec_char)
as.numeric(vec_df$vec_factor)
```

* Logical vector converts to 0 and 1.
* Character vector gives warning:NAs introduced by coercion.
* Factor vecter returns numeric numbers, which is based on the alphabetical orders in the sequence. 


* In a second code chunk, convert your character variable from character to factor to numeric.
similarly, convert your factor variable from factor to character to numeric.

### convert to numeric part2

```{r convert,, eval=FALSE}
as.numeric(as.factor(vec_df$vec_char))
as.numeric(as.character(vec_df$vec_factor))
```

* Converting character variable from character to factor to numeric returns a vector of numeric numbers.
* Converting factor variable from factor to character to numeric returns Warning: NAs introduced by coercion.

# problem 2

Create a data frame comprised of:
x: a random sample of size 1000 from a standard Normal distribution
y: a random sample of size 1000 from a standard Normal distribution
A logical vector indicating whether the x + y > 0
A numeric vector created by coercing the above logical vector
A factor vector created by coercing the above logical vector


### create a data frame

```{r problem2}

set.seed(1)

plot_df = tibble(
  x=rnorm(1000),
  y=rnorm(1000),
  vec_logical= x+y >0,
  vec_numeric =as.numeric(vec_logical),
  vec_factor = as.factor(vec_logical)
  
)
```
* the size of the dataset is `r nrow(plot_df)` rows by `r ncol(plot_df)` columns.
* The mean is `r mean(plot_df$x)`
* The median is  `r median(plot_df$x)`
* the proportion of cases for which the logical vector is true  `r sum(plot_df$vec_logical == TRUE)/length(plot_df$vec_logical) `

### make a scatterplot

* Make a scatterplot of y vs x; color points using the logical variable 

```{r plot}
library(tidyverse)

ggplot(plot_df, aes(x = x, y = y, color=vec_logical)) + geom_point()

ggsave("scatter_plot1.pdf")
```

* In this case, blue points indicate TRUE and red points indicate FALSE. Only two colors are in the color scale.

* Make a second and third scatterplot that color points using the numeric and factor variables

```{r plot2}
library(tidyverse)

ggplot(plot_df, aes(x = x, y = y, color=vec_numeric)) + geom_point()

ggplot(plot_df, aes(x = x, y = y, color=vec_factor)) + geom_point()
```

* In the second plot, the color scale uses a gradident. Color changes from bright blue to dark blue. Because there are only two different values in  vec_numeric, only two colors are shown on the plot.

* In the third plot, only two discrete colors are in the color scale. Blue points indicate TRUE and red points indicate FALSE.
